---
title: "Tidyverse"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Tidyverse

In this assignment, you’ll practice collaborating around a code project with GitHub.  You could consider our collective work as building out a book of examples on how to use TidyVerse functions.

GitHub repository:  https://github.com/acatlin/FALL2020TIDYVERSE

FiveThirtyEight.com datasets.

Kaggle datasets. 

Your task here is to Create an Example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. (25 points)


#### Read data using readr::read_csv

```{r }
library(tidyverse)

pums <- read_csv("https://raw.githubusercontent.com/szx868/data607/master/Tidyverse/2019PUMS_PERSON_DATA_NY.csv")


```
#### Using dplyr::select to have column you needed

```{r}
select(pums,c("Age","SEX","Total_Personal_Earnings","Total_Personal_Income"))

```

#### Using dplyr::filter to age

```{r}
filter(pums,Age > 36)

```

###  Summarization

With the select and filter verbs in your arsenal, it is now appropriate to move onto summarize().  And as you will see, Summarize is often utilized with grouby().

```{r}

pums%>%
  #we remove the NA records to make calculations possible
  filter(!is.na(Total_Personal_Income))%>%
  #here we are creating one column that holds the mean of Total_Personal_Income.
  summarise(avg_income = mean(Total_Personal_Income))
```

We can create as many summarise columns as we need
```{r}

pums%>%
  filter(!is.na(Total_Personal_Income))%>%
  #here we are creating one column that holds the mean of Total_Personal_Income.
  summarise(avg_income = mean(Total_Personal_Income), max_ed = max(Educational_Attainment))
```


As mentioned about the groupby() function adds some interesting capabilities.
```{r}

pums%>%
  filter(!is.na(Total_Personal_Income))%>%
  group_by(Age)%>%
  summarise(avg_income = mean(Total_Personal_Income), max_ed = max(Educational_Attainment))
```



